---
title: "Clustering and PCA"
author: "Bao Doquang, Dhwanit Agarwal, Akksay Singh and Shristi Singh"
date: "April 20, 2020"
output: pdf_document
---

I. Overview: 

The data in wine.csv contains information on 11 chemical properties of 6497 different bottles of vinho verde wine from northern Portugal. Additionally, the data contains two other attributes about each wine. These two attributes are color (whether the wine is red or white) and quality (the quality of the wine, as judged on a 1-10 scale by a panel of certified wine snobs). Our objective is to use just the 11 chemical properties (or suitable transformations thereof) of the 6497 different bottles of vinho verde wine to accurately predict the color and quality of the wine using a PCA and a clustering algorithm. Futhermore we need to identify which dimensionality reduction technique makes more sense for this data. 

Loading libraries:- 
```{r}
library(ggplot2)
library(foreach)
library(LICORS)
library(caret)
library(e1071)
```

II. Data

Loading data-
```{r}
wine <- read.csv("~/GitHub/SDS323_Spring2020/hw3/q3/wine.csv")
str(wine)
```

We take the subset of the data so that only the chemical properties of the wine are in the data set
```{r}
wine_chem_prop <- wine[1:11]
```

We normalize the data by adjusting the scale so that measuring distance between points remains meaningful
```{r}
wine_chem_prop_norm = scale(wine_chem_prop, center = TRUE, scale = TRUE)
```

III. Model & Results

Clustering algorithm to determine color of wine:

We begin with running k-means with 2 clusters and 25 starts and then examine our clusters. 
```{r}
clustk2 = kmeanspp(wine_chem_prop_norm, 2, nstart = 25)

which(clustk2$cluster ==1)
which(clustk2$cluster ==2)
sum(clustk2$withinss)
```

By analyzing the correlation between each chemical property and a given color of wine We try to narrow down the variables that we want to compare before visualizing the clusters. 
```{r}
cor(wine_chem_prop, as.numeric(wine$color))
```
From the output above, We will choose the covariates with the greatest positive or negative correlation with wine color. 
Hence we will choose volatile.acidity, chlorides, total.sulfur.dioxide, and sulphates. In this output, higher positive correlation means that that specific chemical property is more closelyassociated with white wine.

We now compare select variables and wine color:
```{r}
pairs(wine[, c(2,5,7,10)], col = c("red", "grey")[wine$color],
      main = "Chemical Properties of wine: Red vs White")
```

Now, we examine the cluster 
```{r}
pairs(wine[, c(2,5,7,10)], col = c("red", "grey")[clustk2$cluster],
      main = "Chemical Properties of wine: Cluster 1 vs Cluster 2")
```

From a first glance at the graphs it appears that the clustering alogrithm effectively seperates the red from the white wines.

Now lets assign red wine to cluster 1 and white wine to cluster 2, and calculate the confusion matrix.
```{r}
yhat1 = clustk2$cluster
yhat1 = ifelse(yhat1 ==1, "red", "white")

confusionMatrix(data = as.factor(yhat1), reference = wine$color)

```

For comparison, we run the null model.
```{r}
confusionMatrix(data = as.factor(rep("white", length(clustk2$cluster))), reference = wine$color)
```

From looking at the confusion matrix it is apparent that clustering is an effective way of grouping together the red and white wines. The only draw back of this method is that we can not tell exactly which chemical property is associated with which color of wine. 

Principle Components Analysis Algorithm:

We begin by trying the dimension reduction technique, principle components analysis (PCA). PCA allows for mixed membership of covariates to construct principle components. We begin our analysis by trying to classify wine by color.
```{r}
pc_chem = prcomp(wine_chem_prop_norm)
pc_chem$rotation[,1]
```

Data Visualization: 
```{r}
qplot(pc_chem$x[,1], pc_chem$x[,2], 
      color=wine$color, xlab='Component 1', 
      ylab='Component 2',geom = c("point", "abline"),intercept = 2.5, slope = 2)
```

In the graph above we notice that red and white wine split roughly along the line cp2 = cp1 - 2.5. 
Hence, if a point lands on the right side of the line then it is most likely white wine but if the point lands on the left side of the line then it is most likely red wine. 
```{r}

y_hat2 = pc_chem$x[,2] - 2*pc_chem$x[,1] - 2.5
y_hat2 = ifelse(y_hat2 > 0, "red", "white")

confusionMatrix(data = as.factor(y_hat2), reference = wine$color)

```

This method seams fairly accurate. We can get a better prediction if we try to make more fine adjustments to the line but from looking at only the unsupervised informatin we are not sure how much the prediction will get better.
Even the way we defined y_hat2 above required us to peek at that shape of the data.

Next, we try to classify wine by quality. 

We begin by setting k = 10 as wines are rated on a scale of 1 to 10.
We prceed by constructing an elbow plot to check if k = 10 is an appropriate choice for k. 
```{r}

k_grid = seq(2,20, by = 1) # This is a vector containing various values of k.

SSE_grid = foreach(k = k_grid, .combine = 'c') %do% {
  cluster_k = kmeans(wine_chem_prop_norm, k, nstart = 50)
  cluster_k$tot.withinss
}

plot(k_grid, SSE_grid)
```

From the elbow plot above, it is apparent that many k values can be called the elbow and k=10 is a resonable choice among them so we can move forword and cluster the data. 
```{r}
clustk10 = kmeanspp(wine_chem_prop_norm, 10, nstart = 25)

cor(wine_chem_prop, as.numeric(wine$quality))

```

Unfortunately, there are not that many chemical properties with a particular strong corollation with quality.
Some chemcial properties that we select to find the corrolation between that chemical properties and quality are 
volatile.acidity, chlorides, density, and alcohol.    

Now we will go ahead and make a set of pairwise graph of these select chemical properties and fill it by quality score. 
we need to use only 7 colors to fill our points because while the wines could be scored on a scale of 1 to 10 but only scores of 3 through 9 were actually given to the wines in our data. 
```{r}
pairs(wine[, c(2,5,8,11)], 
      col = c("dark red", "red","pink", "grey", "blue", "dark blue", "purple")[wine$quality],
      main = "Chemical Properties of Wine: Quality Scores",
      oma=c(5,5,5,15))
par(xpd = TRUE)
legend("right", 
       fill = c("dark red", "red","pink", "grey", "blue", "dark blue", "purple"), 
       legend = c( levels(as.factor(wine$quality))),
       title = "Score")

```

From the graphs above it is apparent that there are no clear clusters. 

Since this was not that useful we will not try categorizing the wines into only two quality groups, "low quality" wine and "high quality" wine. 
We will define scores below 7 as "low quality" and scores above 7 as "high quality." 
```{r}
wine_qual = wine
wine_qual$quality = ifelse(wine$quality < 7, "low", "high")

pairs(wine_qual[, c(2,5,8,11)], 
      col = c("red","grey")[as.factor(wine_qual$quality)],
      main = "Chemical properties of Wine: High vs Low Quality")
legend("right", 
       fill = c("red","grey"), 
       legend = c( levels(wine$color)))

pairs(wine_qual[, c(2,5,8,11)], 
      col = c("grey","red")[clustk2$cluster],
      main = "Chemical properties of Wine: Cluster 1 vs Cluster 2")

```

We make a confusion matrix, to check if the clustering algorithm preforms significantly better if we only have the two categories low and high quality wines respectively instead of k = 10 many clusters.
```{r}

y_hat3 = clustk2$cluster
y_hat3 = ifelse(clustk2$cluster == 1, "low", "high")

confusionMatrix(data = as.factor(y_hat3), reference = as.factor(wine_qual$quality))
```

Our accuracy is 61.67%, which is worse than our null model.
To see if PCA is more appropriate for this classificiation problem we can use our PCA algorithm results to make our desired plots.
```{r}

qplot(pc_chem$x[,1], pc_chem$x[,2], 
      color=wine$quality, xlab='Component 1', 
      ylab='Component 2')

```

We look at some more pair-wise comparision to be more meticulous.
```{r}

pairs(pc_chem$x[, c(1,2,3,4)],
      col = c("dark red", "red", "blue", "dark blue")[wine$quality])

```

From looking at the results above there is no apparent way to tell which chemical properties correspond to quallity ratings. 

Now, we look on the basis of high and low quality wines that we defined previously on the basis of quality rating.
```{r}

qplot(pc_chem$x[,1], pc_chem$x[,2], 
      color= factor(wine_qual$quality), xlab='Component 1', 
      ylab='Component 2',geom = c("point", "abline"))

```

It seems that there is no visual divide among the types of wine.

To further verify this, we look at pairwise plots for the first four principle components.
```{r}
pairs(pc_chem$x[, c(1,2,3,4)],
      col = c("red", "blue")[as.factor(wine_qual$quality)])
```

Hence, in this case PCA is not capable of sorting the wines on the basis of quality.